---
title: "Times series forecasting"
output: pdf_document
---

#load the packages
```{r}
library(forecast)
library(ggplot2)
```
#set Working Directory
```{r}
getwd()
setwd("/Users/dabasn/Desktop/Nitesh/Studies/DSTI/TimeSeries/TSLabs/Day1/Varicella_files")
```
#load the data and plot them
```{r, out.width='70%'}
data=read.csv('varicelle.csv', header = TRUE)
varicelle<-ts(data$x,start=c(1931,1),end=c(1972,6),freq=12)
plot(varicelle)
```
It seems to be a seasonal pattern. We can check this with the seasonplot

The mean is given by
```{r}
mean(varicelle)
```
And the auto-correlation mean that there is a seasonal pattern in the data
```{r, out.width='70%'}
tmp=acf(varicelle,type="cor",plot = FALSE)
tmp$acf[1:3,1,1]
plot(tmp)
```
What is confirmed by the seasonal plot

```{r, out.width='70%', warning=FALSE}
ggseasonplot(varicelle,year.labels= TRUE,year.labels.left=TRUE)
```
We can also plot manually the seasonal plot

```{r, out.width='70%'}
plot(varicelle[1:12],type="l",ylim=c(min(varicelle),max(varicelle)))
for (i in 1:41) lines(varicelle[(1+12*i):(12*(i+1))])
```
We now compute and plot the annual evolution - To removed the seasonality (Only trend plot)
```{r, out.width='70%'}
x=rep(0,41)
for (i in 0:40) x[i+1]<-sum(varicelle[(1+12*i):(12*(i+1))])
plot(x,type='l',xaxt='n',xlab='')
axis(1,at = 0:40,labels = 1931:1971)
```
Split dataset into training and testing sets and forcast. 
Red - Actual data
Green - Predicted data
```{r}
vari_train <- ts(data$x[1:480],start=c(1931,1),end=c(1970,12),freq=12) 
vari_test <- ts(data$x[481:498],start=c(1971,1),end=c(1972,6),freq=12)
plot(vari_train,xlim=c(1931,1973))
lines(vari_test,col=2)
SES=HoltWinters(vari_train,alpha=NULL,beta=FALSE,gamma=FALSE) 
p1<-predict(SES,n.ahead=18)
lines(p1,col=3)
```
Forecasting with SES
```{r}
SES=ses(vari_train,h=18) 
round(accuracy(SES),2)
autoplot(SES)
```
#Varicella forecasting with non seasonal HW smoothing
```{r}
HOLT=holt(vari_train,h=18)
round(accuracy(HOLT),2)
autoplot(vari_train) + autolayer(HOLT,series='fitted',PI=FALSE) #PI=FALSE remove the prediction interval
```
#Varicella forecasting with non seasonal HW smoothing; 0 < φ < 1
```{r}
HOLT1=holt(vari_train,h=18) 
HOLT2=holt(vari_train,damped=TRUE,phi=0.9,h=18)
autoplot(vari_train) + autolayer(HOLT1,series='HW',PI=FALSE) + autolayer(HOLT2,series='Damped HW',PI=FALSE)
```
#Varicella forecasting with seasonal HW smoothing - Additive OR Multiplicative
```{r}
fit1=hw(vari_train,seasonal='additive',h=18) 
fit2=hw(vari_train,seasonal='multiplicative',h=18) 
fit3=hw(vari_train,seasonal='additive',damped=TRUE,h=18)
fit4=hw(vari_train,seasonal='multiplicative',damped=TRUE,h=18)

autoplot(vari_train) + 
  autolayer(fit1,series='HW add.',PI=FALSE) + 
  autolayer(fit2,series='HW mult.',PI=FALSE)

autoplot(vari_train) + 
  autolayer(fit3,series='HW mult.',PI=FALSE, col='blue') +
  autolayer(fit4,series='HW mult.',PI=FALSE, col='green')
```
#Additive/Multiplicative + Damped versions
#damped version of the seasonal HW
```{r}
#Model Comparision
print(sqrt(mean((fit1$mean-vari_test)^2)))
print(sqrt(mean((fit2$mean-vari_test)^2))) #best model
print(sqrt(mean((fit3$mean-vari_test)^2)))
print(sqrt(mean((fit4$mean-vari_test)^2)))
```

